
\begin{thebibliography}{1}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alphabetical by last name!
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\bibitem{now}
D. Abadi, P. Boncz, S. Harizopoulos, S. Idreos, and Samual Madden.
The Design and Implmentation of Modern Column-Oriented Database Systems.
In Foundations and Trends in Databases, Vol. 5. No. 3, 2012.


\bibitem{columnOverview}
D. Abadi, P. Boncz, and S. Harizopoulos
Column-oriented Database Systems
In VLDB, 2009


\bibitem{slides}
D. Dewitt.
From 1 to 1000 mips
PASS Summit 2009 Keynote.


\bibitem{compression}
D. J. Abadi, S. R. Madden, N. Hachem.
Integrating Compression and Execution in Column-Oriented Database Systems.
In SIGMOD, 2006.


\bibitem{colvsrow}
D. J. Abadi, S. R. Madden, N. Hachem.
Column-Stores vs. Row-Stores: How Different Are They Really?
In SIGMOD, 2008.


\bibitem{materialization}
D. J. Abadi, D. S. Myers, D. J. DeWitt, and S. R. Madden.
Materialization Strategies in a Column-Oriented DBMS.
In ICDE, 2007.


\bibitem{dbvm}
A. Aboulnaga, U. Minhas, K. Salem, J. Yadav.
Database Systems on Virtual Machines: How Much do You Lose? 
In ICDE Workshops, pages 35-41, 2008.


\bibitem{dbaasazure}
T. Cabot and L. Pizette.
Database as a Service: A Marketplace Assessment.
Web Page, Jan. 2012.
http://www.mitre.org/sites/default/files/pdf/cloud\_database\_service\_dbaas.pdf


\bibitem{dbaas}
V. Cheung.
DBaaS 101: What It Is and What to Expect.
Web page, March 2014.
http://www.geniedb.com/dbaas-101-what-it-is-and-what-to-expect/


\bibitem{relcloud}
C. Curino, E. Jones, R. Popa, N. Malviya, E. Wu, S. Madden, H. Balakrishnan, N. Zeldovich.
Relational Cloud: A Database-as-a-Service for the Cloud.
In CIDR, paper 33, 2011.


\bibitem{dynamo}
G. DeCandia, D. Hastorun, M. Jampani, G. Kakulapati, A. Lakshman, A. Pilchin, S. Sivasubramanian, P. Vosshall and W. Vogels.
Dynamo: Amazon's Highly Available Key-value Store.
SOSP, 2007.


\bibitem{sql}
P. Larson, E. N. Hanson, and S. L. Price.
Columnar Storage in SQL Server 2012
In IEEE bullitin, 2012.


\bibitem{dbseer}
B. Mozafari, C. Curino and S. Madden.
DBSeer: Resource and Performance Prediction for Building a Next Generation Database Cloud.
In CIDR, paper 52, 2013.

\bibitem{cstore}
D. Abadi, P. Boncz, S. Harizopoulos, S. Idreos, and Samual Madden.
C-Store: A Column-Oriented DBMS
Web Page: http://db.csail.mit.edu/projects/cstore/


\bibitem{starschema}
P. O'Neil, B. O'Neil, X. Chen.
Star Schema Benchmark.
Web Page (Revision 3, 2009): http://www.cs.umb.edu/~poneil/StarSchemaB.PDF.


\bibitem{vertica}
Vertica.
Vertica Crunchbase.
Web Page: http://www.crunchbase.com/company/vertica-systems.


\bibitem{CPU}
H. Plattner.
A Common Database Approach for OLTP and OLAP Using and In-Memory Column Database.
In SIGMOD, 2009.

\bibitem{monet}
S. Ideros, F. Groffen, N. Nes, S. Manegold, S. Mullender, M. Kersten.
MonetDB: Two Decades of Research in Column-oriented Database Architectures.
IEEE Data Eng. Bull., 35(1):40-45, 2012.


\bibitem{pattern_match}
L. Zou, L. Chen, M. Tamer Ozsu.
Distance-Join: Pattern Match Query in a Large Graph Database
In VLDB, 2009.

\bibitem{graphsys}
Y. Lu, J. Cheng, D. Yan, H. Wu.
Large-Scale Distributed Graph Computing Systems: An Experimental Evaluation.
In VLDB, 2014.

\bibitem{blogel}
D.Yan, J. Cheng, W. Ng.
Blogel: A Block-Centric Framework for Distributed Computation on Real-World Graphs.
In VLDB, 2014.

\bibitem{pregel}
G. Malewicz, M. Austern, A. Bik, J. Dehnert, I. Horn, N. Leiser, G. Czajkowski.
Pregel: A System for Large-Scale Graph Processing.
In SIGMOD, 2010.

\bibitem{RDF}
T. Neumann, G. Weikum.
RDF-3X: a RISC-style Engine for RDF.
In PVLDB, 2008.

\bibitem{triplestore}
J. Rusher.
Triple Store.
In W3C, 2003.

\bibitem{parallelgraph}
A. Lumsdaine, D. Gregor.
Challenges in Parallel Graph Processing.
World Scientific Publishing Company.

\bibitem{algos}
N. Satish, N. Sundaram, M. M. A. Patwary, J. Seo, J. Park, M. A. Hassaan, S. Sengupta, Z. Yin, and P. Dubey. Navigating the maze of graph analytics frameworks using massive graph datasets. 
In SIGMOD Conference, pages 979–990, 2014.

\bibitem{RDFvsGraph}
D. De Abreu, A. Flores, G. Palma, V. Pestana, J. Pinero, J. Queipo, J. Sanchez, M. Vidal.
Choosing Between Graph Databases and RDF Engines for Consuming and Mining Linked Data.
Universidad Simon Bolivar, 2013.
%http://ceur-ws.org/Vol-1034/DeAbreuEtAl_COLD2013.pdf

\bibitem{scalableRDF}
J. Huang, D. Abadi, K. Ren.
Scalable SPARQL Querying of Large RDF Graphs
In VLDB, 2011.
%http://cs-www.cs.yale.edu/homes/dna/papers/sw-graph-scale.pdf

\bibitem{monetcpu}
P. Boncz, M. Zukowski, N. Nes.
MonetDB/X100: Hyper-Pipelining Query Execution.
In VLDB, 2005.
%http://www.cidrdb.org/cidr2005/papers/P19.pdf

\bibitem{vectorwise}
M. Zukowski, M. van de Wiel, P. Boncz.
Vectorwise: a Vectorized Analytical DBMS.
Actian, 2013.
%http://oai.cwi.nl/oai/asset/19958/19958B.pdf

\bibitem{PAX}
A. Ailamaki, D. DeWitt, M. Hill.
Data Page Layouts for Relational Databases on Deep Memory Hierarchies.
In VLDB, 2002.
%http://research.cs.wisc.edu/multifacet/papers/vldbj02_pax.pdf

\bibitem{redshift}
Amazon Redshift API
Web Page: http://docs.aws.amazon.com/redshift/
Amazon, 2012.

\bibitem{inmemory}
H. Plattner.
The Impact of Columnar In-Memory Databases on Enterprise Systems.
In VLDB, 2014.

\bibitem{kxsystems}
kdb+ Tutorial.
Web Page: http://www.timestored.com/kdb-guides/
Kx Systems.

\bibitem{semantic}
S. Decker, F. van Harmelen, J. Broekstra, M. Erdmann, D. Fensel, I. Horrocks 2, M. Klein, S. Melnik.
The Semantic Web - on the respective Roles of XML and RDF.

\bibitem{google}
S. Brin, L. Page.
The Anatomy of a Large-Scale Hypertextual Web Search Engine.
Stanford University.
%http://infolab.stanford.edu/pub/papers/google.pdf

\bibitem{BFS}
H. Shang, M. Kitsuregawa.
Efficient Breadth-First Search on Large Graphs with
Skewed Degree Distributions.
EDBT/ICDT, 2013.
%http://www.edbt.org/Proceedings/2013-Genova/papers/edbt/a29-shang.pdf

\bibitem{triangles}
M. Kolountzakis, G. Miller, R. Peng, C. Tsourakakis.
Efficient Triangle Counting in Large Graphs via Degree-based Vertex
Partitioning.
In Workshop on Algorithms and Models for the Web Graph, 2010.
%http://people.seas.harvard.edu/~babis/int-math-triangles.pdf




\end{thebibliography}


